---
title: 'Visualizing Defaulted Loans'
---


```{r, echo=F}
options(scipen = 6)
knitr::opts_chunk$set(cache=FALSE, fig.height=3, fig.width = 7, comment=NULL, eval=T, tidy=F, width=80, message= FALSE, echo= FALSE, warning = FALSE)
```

```{r, child="_preamble.Rmd"}
```


```{r, child="_nofilter.Rmd"}
```

## Plotting Defaults by Time to the Last Payment   

To start understanding the defaults, I plotted the relationship between the time to default and the amount of principal repaid.  The x axis is the number of days between the last payment and the day the loan was issued. The y axis the percent of loans that have defaulted.

The relationship should be linear.  You'd expect the lender to consistenly payback principle as they make payments until they eventually default. The interesting part of the chart, then, is how far up the curve the lender was before making its last payment. 

```{r}
# plot1<- LoanData %>%
#     filter(loan_status == "Charged Off") %>%
#     mutate(PR = total_rec_prncp/funded_amnt,
#            DaysToDefault = as.Date(last_pymnt_d)-as.Date(issue_d)) %>%
#     select(PR, DaysToDefault, term) %>%
#     ggplot +
#     aes(x=DaysToDefault, y=PR, group=term, color= term) +
#     geom_point(size=.5, alpha=.05) +
#     facet_grid(term~., scales="free")+
#     geom_smooth() +
#     theme_LC() +
#     labs(x="Days To Last Payment",
#          y="Payback Ratio")

LoanData %>%
    filter(loan_status == "Charged Off") %>%
    mutate(DaysToDefault = as.Date(last_pymnt_d)-as.Date(issue_d)) %>%
	group_by(term, DaysToDefault) %>%
	summarise(N=n()) %>%
	mutate(cum=cumsum(N),
		   Pct= cum/sum(N)) %>%
    ggplot +
    aes(x=DaysToDefault, y=Pct, group=term, color= term) +
    geom_line() +
    theme_LC() +
    labs(x="Days Until Last Payment",
         y="Cumulative loans defaulted")

```

Taking a closer look at the 3 year loans, I've broken out the cumulative defaults by grade.  As expected, the lower grade loans (G) are defaulting faster than the higher grade (A). One interesting observatin is that the grades move in pairs. The paths A and B grade loans take are very similar.  The paths of C and D grade loans are very similar and the same again for E & F grade loans.  If these pairs have similar default paths but different interest rates, then investing in the lower grade note could be a form of arbitration.  

```{r}
LoanData %>%
    filter(loan_status == "Charged Off",
    	   term ==36) %>%
    mutate(DaysToDefault = as.Date(last_pymnt_d)-as.Date(issue_d)) %>%
	group_by(term, grade, DaysToDefault) %>%
	summarise(N=n()) %>%
	mutate(cum=cumsum(N),
		   Pct= cum/sum(N)) %>%
    ggplot +
    aes(x=DaysToDefault, y=Pct, group=grade, color= grade) +
    geom_line() +
	# facet_wrap(~term) +
    theme_LC() +
    labs(x="Days Until Last Payment",
         y="Cumulative loans defaulted")


```

