---
title: "Loan Amount"
output: html_document
vignette: >
    %\VignetteIndexEntry{Exploratory Analysis- Loan Amount}
    %\VignetteEngine{knitr::rmarkdown}
    \usepackage[utf8]{inputenc}
---

```{r, echo=F}
options(scipen = 6)
knitr::opts_chunk$set(cache=FALSE, fig.height=3, fig.width = 7, comment=NULL, eval=T, tidy=F, width=80, message= FALSE, echo= FALSE, warning = FALSE)
```

```{r, child="snippet/_preamble.Rmd"}
```

### Analyzing the distribution by grade

**Quantity by status and grade ** 
```{r}
knitr::kable(format(table(LoanData$Class, LoanData$grade),big.mark=",", trim=TRUE), align='r')
```

**Proportion by status and grade**

```{r}
knitr::kable(round(prop.table(table(LoanData$Class, LoanData$grade),2),2))
```
<hr>
### Analyzing the Loan Amount
```{r}
g<- ggplot(LoanData)
g<- g + aes(loan_amnt, group=grade, colour=grade)
g<- g + geom_density(size= 1)
g<- g + labs(title= "Distribution of loan amount by grades",
             x= "Loan Amount")
g<- g + theme_LC()
g
```

**QUESTION: Why is there a peak at 24,000 and not at 25,000? IS there a cutoff in the scoring model or is 25K a break point for something else?**

```{r, fig.height=6}
g<- ggplot(LoanData)
g<- g + aes(loan_amnt, group=Class, colour=Class)
g<- g + geom_density(size=.75)
g<- g + labs(x= "Loan Amount")
g<- g + facet_grid(grade~.)
g<- g + theme_LC()
g<- g + labs(title= "Distribution of loan amount by default status")
g
```

Looking at grades D through G, the distribution of the defaulted and is fairly consitent with the distribution of current loans. On it's own, this suggests that loan amount won't have a predictive value over default rates.  The exception is there seems to be a slight bump at the $35,000 level which is the maximum Lending Club offers. Loans at the maximum amount default at a disproportionately low rate.  There may be an advantage in investing in loans in these grades at the max loan amount.

There isn't the same hump at the $35,000 level when looking at grade A & B loans.  Those loans are more highly concentrated at the lower loan amount. Perhaps LC's credit rating status punishes the borrower for a higher loan amount. 


**QUESTION: Why does the 35,000 loan amount have lower proportionate defaults in grades C though E? Take a look at 35K loans and the relationship with other variables**


```{r, fig.height=6}
g<- ggplot(LoanData)
g<- g + aes(loan_amnt, group=Class, colour=Class)
g<- g + geom_density(size=.75)
g<- g + labs(x= "Loan Amount")
g<- g + facet_grid(format(LoanData[,"issue_d"],'%Y')~.)
g<- g + theme_LC()
g<- g + labs(title= "Distribution of loan amount by vintage")
g
```

#### Package Version
The analysis was based on the flowing version of the LendingClubData.

```{r}
packageVersion("LendingClubData")
```
